Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-04-29 concurrency/CONCUR 2023/paper.tex
finished proof mechanization part

Recall that in the lock-coupling pattern, each thread always holds at least one lock while traversing the data structure, ensuring that other threads cannot interfere with the traversal or invalidate the ongoing search. Consequently, before calling the \texttt{traverse} function, a thread must acquire the lock for the current node $\texttt{n}$ and obtain the resources $\mathsf{R}(\texttt{n})$ guarded by the lock. These resources may depend on the specific data structure, but always include a piece of \emph{ghost state} describing the current state of the node (its key, value, and key range) shared between the invariant $\mathsf{R}$ and the abstract state $\treerep$, ensuring that the lock and the abstract state agree on the contents of the node. Formally, the resource  $\mathsf{R}(\texttt{n}) \triangleq \mathsf{ghost\_node(\texttt{n}, k, v, range)} \ \ast \ \mathsf{DS\_specific(\texttt{n})}$.

@@

\section{Proof Mechanization}
\label{proof_mech}
All the proofs described above have been mechanized in VST, using its extensions for logical atomicity~\cite{iris-vst-arxiv}. Statistics on the verification effort are shown in Table~\ref{table_coq}

@@

%By utilizing VST, we can create programs in standard C, automatically generate abstract syntax trees (ASTs) for them in Coq, and subsequently state and interactively prove specifications for these programs with the support of automated tactics. Table \ref{table_coq} presents a summary of our Coq development. We formalized all previously described proofs in Coq, taking advantage of VST's features for verifying separation logic specifications of C functions. This encompasses supplementary proofs, template specifications (encompassing all helper functions such as \texttt{findNext}, \texttt{insertOp}, and others, in addition to \texttt{traverse}, \texttt{insert}, \texttt{lookup}, and \texttt{delete}), as well as proofs against with lock-coupling and give-up template specifications. 
\wm{Rather than re-stating what we proved, this is the place to discuss the development: e.g., which parts took the most effort and LoC, which parts we expect to be reusable and in what way, how much bigger or smaller the template proofs were than non-template versions, etc.}

The proofs can be found online at

\href{https://github.com/PrincetonUniversity/DeepSpecDB/tree/master/concurrency/templates}{\color{blue}{https://github.com/PrincetonUniversity/DeepSpecDB/tree/master/concurrency/templates}}

>>>>

Recall that in the lock-coupling pattern, each thread always holds at least one lock while traversing the data structure, ensuring that other threads cannot interfere with the traversal or invalidate the ongoing search. Consequently, before calling the \texttt{traverse} function, a thread must acquire the lock for the current node $\texttt{n}$ and obtain the resources $\mathsf{R}(\texttt{n})$ guarded by the lock. These resources may depend on the specific data structure, but always include a piece of \emph{ghost state} describing the current state of the node (its key, value, and key range) shared between the invariant $\mathsf{R}$ and the abstract state $\treerep$, ensuring that the lock and the abstract state agree on the contents of the node. Formally, the resource  $\mathsf{R}(\texttt{n}) \triangleq \mathsf{ghost\_node(\texttt{n}, k, v, range)} \ \ast \ \mathsf{DataStruct\_specific(\texttt{n})}$.

@@

\section{Proof Mechanization}
\label{proof_mech}
All the proofs described above have been mechanized in VST, using its extensions for logical atomicity~\cite{iris-vst-arxiv}. Statistics on the verification effort are shown in Table~\ref{table_coq}.

@@

%By utilizing VST, we can create programs in standard C, automatically generate abstract syntax trees (ASTs) for them in Coq, and subsequently state and interactively prove specifications for these programs with the support of automated tactics. Table \ref{table_coq} presents a summary of our Coq development. We formalized all previously described proofs in Coq, taking advantage of VST's features for verifying separation logic specifications of C functions. This encompasses supplementary proofs, template specifications (encompassing all helper functions such as \texttt{findNext}, \texttt{insertOp}, and others, in addition to \texttt{traverse}, \texttt{insert}, \texttt{lookup}, and \texttt{delete}), as well as proofs against with lock-coupling and give-up template specifications. 
%\wm{Rather than re-stating what we proved, this is the place to discuss the development: e.g., which parts took the most effort and LoC, which parts we expect to be reusable and in what way, how much bigger or smaller the template proofs were than non-template versions, etc.}
The construction of proofs for the \texttt{traverse} function, in the contexts of both lock-coupling and give-up templates, demanded substantial effort and lines of code (LoC). However, we expect that functions such as \texttt{traverse}, \texttt{inRange}, \texttt{insert}, and \texttt{lookup} will be reusable with only minor adjustments needed.  The methods \texttt{findNext} and \texttt{insertOp} may also be adapted to other data structures, such as linked lists, with some additional effort. 
Specifically, when attempting to prove concurrent linked lists, we plan to tweak the specifications of \texttt{findNext} and \texttt{insertOp} to restructure the \texttt{node} struct. While the size of our template proofs currently approximates that of the no-template versions, we anticipate that strategic application of simple tactics for quantifier expansion could significantly streamline these proofs in the future. The full development of our mechanization effort is available online
\href{https://github.com/PrincetonUniversity/DeepSpecDB/tree/master/concurrency/templates}{\color{blue}{https://github.com/PrincetonUniversity/DeepSpecDB/tree/master/concurrency/templates}}.
